Improving Coconut Rhinoceros Beetle Breeding Site Detection Using Harmonic Radar

### 2022-02-16 FS-CRB-HR-report3.tex
report3

\begin{document}
\titlehead{US Forest Service Forest Health Protetion Grant Progess Report 2}
\title{Improving Coconut Rhinoceros Beetle Breeding Site Detection Using Harmonic Radar}
\author{Aubrey Moore, University of Guam}
%\date{Submitted December 30, 2020\\Revised January 28, 2021}
\maketitle
\begin{description}	
	\item[GRANTEE:] Aubrey Moore, University of Guam 
	\item[GRANT YEAR:] 2020
	\item[GRANT NUMBER:] 20-DG-11052021-227
	\item[GRANT PROGRAM:] Forest Health Protection
	\item[GRANT EXPIRATION DATE:] 2022-12-31 (extended)
	\item[DATES COVERED BY THIS REPORT:] 2021-01-01 through 2021-06-01
	\item[GRANT STATUS:] Active
\end{description}

@@

\begin{footnotesize}
\url{https://github.com/aubreymoore/Harmonic-Radar/raw/master/FS-CRB-HR-report2.pdf}
\end{footnotesize}

@@

\textbf{The RECCO hand-held harmonic radar device we are using is designed for ground-based location of tags. We are collaborating with Dr. Glenn Dulla, Guam Department of Agriculture, to test the idea of mounting the device on a drone to allow efficient downward-looking scans over large areas. The idea is to mount the RECCO device on a drone which is programmed to fly a defined search path. A digital recorder cabled to the RECCO's headphone jack records the audio signal from the device and the drone's GPS  records location. Georeferencing is done by matching timestamps in the audio recording with timestamps in the GPS data. Postprocessing generates a map of signal strength with peaks indicating probable tag locations requiring confirmation by ground searches.
}

\textbf{Preliminary tests are promising. To date we have:
\begin{itemize}
	\item Used digital recordings of the RECCO's audio output to locate tags (See Technical Report: Using a Cell Phone as a Datalogger for Handheld Harmonic Radar \url{https://github.com/aubreymoore/Harmonic-Radar/blob/master/experiments/techreport/techreport.pdf})	
	\item Flown the RECCO harmonic radar device attached to a drone (Fig. \ref{fig:drone})
\end{itemize}
}

@@

\section{MONITORING \& EVALUATION}

@@

\begin{tabular}{lrrl}
	\hline
	Category & Budget & Spent & Note \\
	\hline 
	Equipment and supplies & \$8,000 & \$3,462 & Harmonic radar and tags \\ 
	Travel & \$12,000 & \$0 \\ 
	Admin. fee & \$3,000 & \$0 \\ 
	\hline 
\end{tabular}

@@

\textbf{COVID travel restrictions during the current reporting period did not allow Dr. Siderhurst and students to visit Guam to perform the planned field work. Guam is currently experiencing high COVID incidence predominantly caused by the D variant. We are currently targeting spring break 2022 as the time period for field work.}

>>>>

\begin{document}
\titlehead{US Forest Service Forest Health Protetion Grant Progess Report 3}
\title{Improving Coconut Rhinoceros Beetle Breeding Site Detection Using Harmonic Radar}
\author{Aubrey Moore, University of Guam}
%\date{Submitted December 30, 2020\\Revised January 28, 2021}
\maketitle
\begin{description}	
	\item[GRANTEE:] Aubrey Moore, University of Guam 
	\item[GRANT YEAR:] 2020
	\item[GRANT NUMBER:] 20-DG-11052021-227
	\item[GRANT PROGRAM:] Forest Health Protection
	\item[GRANT EXPIRATION DATE:] 2022-12-31 (extended)
	\item[DATES COVERED BY THIS REPORT:] 2021-06-01 through 2021-12-31
	\item[GRANT STATUS:] Active
\end{description}

@@

\begin{footnotesize}
\url{https://github.com/aubreymoore/Harmonic-Radar/raw/master/FS-CRB-HR-report3.pdf}
\end{footnotesize}

@@

The RECCO hand-held harmonic radar device we are using is designed for ground-based location of tags. We are collaborating with Dr. Glenn Dulla, Guam Department of Agriculture, to test the idea of mounting the device on a drone to allow efficient downward-looking scans over large areas. The idea is to mount the RECCO device on a drone which is programmed to fly a defined search path. A digital recorder cabled to the RECCO's headphone jack records the audio signal from the device and the drone's GPS  records location. Georeferencing is done by matching timestamps in the audio recording with timestamps in the GPS data. Postprocessing generates a map of signal strength with peaks indicating probable tag locations requiring confirmation by ground searches.

Preliminary tests are promising:
\begin{itemize}
	\item Used digital recordings of the RECCO's audio output to locate tags (See Technical Report: Using a Cell Phone as a Datalogger for Handheld Harmonic Radar \url{https://github.com/aubreymoore/Harmonic-Radar/blob/master/experiments/techreport/techreport.pdf})	
	\item Flown the RECCO harmonic radar device attached to a drone (Fig. \ref{fig:drone})
	\item \textbf{Based on initial flight tests we determined that the audio data recorder we were using was too heavy for use on the drone. We purchased a much smaller and light audio data recorder. Walk tests with this recorder show that we can locate harmonic radar tags (Fig. \ref{fig:211109003}).}
\end{itemize}

\begin{figure}[h]
	\centering
	\includegraphics[width=\linewidth]{211109_003.png}
	\caption{Audio output from a RECCO harmonic radar device recorded using a Zoom F2 field recorder. This was a walk test along a linear transect in which a harmonic radar tag had been placed. Location of the tag in the recording is clear.}
	\label{fig:211109003}
\end{figure}

@@

\textbf{On December 18, Aubrey Moore made a presentation on the technology used in this project at a meeting of the Guam Beekeepers Association (GBA). Members of the GBA and the Guam Department of Agriculture are interested in using harmonic radar to track greater banded hornets (GBH), \textit{Vespa tropica}, to their nests so that they can be destroyed. GBH, a recently arrived invasive species which raids honeybee hives, is a serious threat to Guam's nascent beekeeping industry.
}
\section{MONITORING \& EVALUATION}

@@

\begin{tabular}{lrrl}
	\hline
	Category & Budget & Spent & Note \\
	\hline 
	Equipment and supplies & \$8,000 & \$316.93 & Miniturized audio field recorder and hardware \\ 
	Travel & \$12,000 & \$0 \\ 
	Admin. fee & \$3,000 & \$0 \\ 
	\hline 
\end{tabular}

@@

COVID travel restrictions during the current reporting period did not allow Dr. Siderhurst and students to visit Guam to perform the planned field work. Guam is currently experiencing high COVID incidence predominantly caused by the D variant. We are currently targeting spring break 2022 as the time period for field work.

\textbf{As of mid February 2022, Guam is still has extremely high local transmission of COVID due to delayed arrival of the OMICRON variant. However, quarantine for fully vaccinated visitors from the mainland is no longer required. A decision on when Dr. Siderhurst and students experienced with using harmonic radar to track insects will visit Guam to complete planned field work is still pending.}
